---
title: "MSAN 593 - Homework 2"
author: "Andre Guimaraes Duarte"
date: "July 24, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TODO: reference histograms in the text

# Question 1
## 1.1
### 1.1.1
```{r myRunifVec, fig.cap="Histogram of 10,000,000 random variables ~ U(4, 6)\\label{runifPop}"}
times <- 10000000
myRunifVec <- runif(times, 4, 6)
hist(myRunifVec,
     main = paste("Histogram of ", length(myRunifVec), "\n random variables ~ U(4, 6)",
                  sep = ""), xlab = "x", freq = F)
```

The histogram in figure \ref{runifPop} is the histogram of a $\sim U(4, 6)$ distribution with `r as.integer(length(myRunifVec))` random variables. 

### 1.1.2
```{r sampleUnif, fig.cap="Histogram of a sample of 100,000 random elements of a ~U(4, 6) distribution\\label{runifSample}"}
samples <- 100000
myRunifSample <- sample(myRunifVec, samples)
hist(myRunifSample,
     main = paste("Histogram of ", length(myRunifSample),
                  "\n random variables sampled from a ~ U(4, 6)",
                  sep = ""), xlab = "x", freq = F)
```

The histogram in figure \ref{runifSample} is very similar to the population distribution \ref{runifPop}. Sampling elements from a uniform distribution maintains its original distribution.

### 1.1.3
```{r runifMean2, fig.cap="Histogram of a sample of 100,000 random means of two elements of a ~U(4, 6) distribution\\label{runif2}"}
# How to make it more efficient for 100,000??
n <- 10
unifSampleMean_2 <- replicate(n, {
  mean(sample(myRunifVec, 2))
})
hist(unifSampleMean_2)
```

The histogram in figure \ref{runif2} is not similar to the population distribution shown in figure \ref{runifPop}. In fact, we can start to see a bell-shaped distribution in this figure.

### 1.1.4
```{r runifMean5, fig.cap="Histogram of a sample of 100,000 random means of five elements of a ~U(4, 6) distribution\\label{runif5}"}
unifSampleMean_5 <- replicate(n, {
  mean(sample(myRunifVec, 5))
})
hist(unifSampleMean_5)
```

The histogram in figure \ref{runif5} is different from the population distribution in figure \ref{runifPop}. There are more observations around $5$, and less toward the tails.

### 1.1.5
```{r runifMean10, fig.cap="Histogram of a sample of 100,000 random means of ten elements of a ~U(4, 6) distribution\\label{runif10}"}
unifSampleMean_10 <- replicate(n, {
  mean(sample(myRunifVec, 10))
})
hist(unifSampleMean_10)
```

The histogram in figure \ref{runif10} is different from the population distribution in figure \ref{runifPop}. This one is bell-shaped and seems to be symmetrical around the value $5$.

### 1.1.6
```{r runifMean30, fig.cap="Histogram of a sample of 100,000 random means of thirty elements of a ~U(4, 6) distribution\\label{runif30}"}
unifSampleMean_30 <- replicate(n, {
  mean(sample(myRunifVec, 30))
})
hist(unifSampleMean_30)
```

The histogram in figure \ref{runif30} is different from the population distribution in figure \ref{runifPop}. It looks a lot like a normal distribution centered around $5$.

***

## 1.2
### 1.2.1
```{r myRexpVec, fig.cap="Histogram of 10,000,000 random variables ~Exp(0.5)\\label{rexpPop}"}
myRexpVec <- rexp(times, 0.5)
hist(myRexpVec,
     main = paste("Histogram of ", length(myRexpVec), "\n random variables ~ Exp(0.5)",
                  sep = ""), xlab = "x", freq = F)
```

Figure \ref{rexpPop} shows the histogram of a $\sim Exp(0.5)$ distribution with `r as.integer(length(myRexpVec))` random variables.

### 1.2.2
```{r sampleExp, fig.cap="Histogram of a sample of 100,000 random elements of a ~Exp(0.5) distribution\\label{rexpSample}"}
myRexpSample <- sample(myRexpVec, samples)
hist(myRexpSample,
     main = paste("Histogram of ", length(myRexpSample),
                  "\n random variables sampled from a ~ Exp(0.5)",
                  sep = ""), xlab = "x", freq = F)
```

The histogram in figure \ref{rexpSample} is very similar to the population distribution \ref{rexpPop}. Sampling elements from a negative exponential distribution maintains its original distribution.

### 1.2.3
```{r rexpMean2, fig.cap="Histogram of a sample of 100,000 random means of two elements of a ~Exp(0.5) distribution\\label{rexp2}"}
expSampleMean_2 <- replicate(n, {
  mean(sample(myRexpVec, 2))
})
hist(expSampleMean_2)
```

Figure \ref{rexp2} shows the histogram is slightly more skewed to the right than the population distribution in figure \ref{rexpPop}. 

### 1.2.4
```{r rexpMean5, fig.cap="Histogram of a sample of 100,000 random means of five elements of a ~Exp(0.5) distribution\\label{rexp5}"}
expSampleMean_5 <- replicate(n, {
  mean(sample(myRexpVec, 5))
})
hist(expSampleMean_5)
```

The histogram when we take the mean of five elements from the original population, shown in figure \ref{rexp5}, shows more skewness to the right than the histogram of the population distribution in figure \ref{rexpPop}.

### 1.2.5
```{r rexpMean10, fig.cap="Histogram of a sample of 100,000 random means of ten elements of a ~Exp(0.5) distribution\\label{rexp10}"}
expSampleMean_10 <- replicate(n, {
  mean(sample(myRexpVec, 10))
})
hist(expSampleMean_10)
```

By averaging ten elements, the histogram in figure \ref{rexp10} is starting to look a lot like a Gamma distribution.

### 1.2.6
```{r rexpMean30, fig.cap="Histogram of a sample of 100,000 random means of thirty elements of a ~Exp(0.5) distribution\\label{rexp30}"}
expSampleMean_30 <- replicate(n, {
  mean(sample(myRexpVec, 30))
})
hist(expSampleMean_30)
```

In figure \ref{rexp30}, we see the histogram when we take the average of 30 elements from the original negative exponential population shown in figure \ref{rexpPop}. The distribution is now very different, and looks like a Gamma distribution.

***

## 1.3
### 1.3.1
```{r myBdist, fig.cap="Histogram of 5,000,000 random variables from a ~N(-3, 1) and 5,000,000 random variables from a ~N(3, 1)\\label{bdistPop}"}
times <- 5000000
myBdist <- c(rnorm(times, -3, 1), rnorm(times, 3, 1))
hist(myBdist, main="Bimodal histogram of ~N(-3, 1) and ~N(3, 1)", xlab="x", freq = F)
```




### 1.3.2
```{r bdist5, fig.cap="Histogram of a sample of 100,000 random means of five elements from a ~N(-3, 1) and a ~N(3, 1) distributions\\label{bdist5}"}
myBdist_5 <- replicate(n, {
  mean(sample(myBdist, 5))
})
hist(myBdist_5)
```



### 1.3.3
```{r bdist10, fig.cap="Histogram of a sample of 100,000 random means of ten  elements from a ~N(-3, 1) and a ~N(3, 1) distributions\\label{bdist10}"}
myBdist_10 <- replicate(n, {
  mean(sample(myBdist, 10))
})
hist(myBdist_10)
```

```{r bdist20, fig.cap="Histogram of a sample of 100,000 random means of twenty elements from a ~N(-3, 1) and a ~N(3, 1) distributions\\label{bdist20}"}
myBdist_20 <- replicate(n, {
  mean(sample(myBdist, 20))
})
hist(myBdist_20)
```

```{r bdist30, fig.cap="Histogram of a sample of 100,000 random means of thirty elements from a ~N(-3, 1) and a ~N(3, 1) distributions\\label{bdist30}"}
myBdist_30 <- replicate(n, {
  mean(sample(myBdist, 30))
})
hist(myBdist_30)
```

### 1.3.4








